\documentclass[a4paper, 10pt, oneside, article]{memoir}
\chapterstyle{culver}
\checkandfixthelayout

\usepackage{lipsum}

% Palatino font
\usepackage{palatino}

% Font and input encoding
\usepackage[T1]{fontenc}
\usepackage[utf8x]{inputenc}

% Babel (language)
\usepackage[english]{babel}

% Support for blackboard bold symbols.
\usepackage{bbm}

% AMS-Math packages
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}

% For including bitmap graphics
\usepackage{graphicx}

\usepackage{multirow}

\usepackage{todonotes}

% Remove chapters from figure counters
%\counterwithout{figure}{chapter}
% Add sections instead
%\counterwithin{figure}{section}
%\counterwithout{section}{chapter}

\usepackage{tikz}
\usetikzlibrary{shapes,arrows}

\title{Advanced Algorithms and Data Structures\\ Branch and Bound}
\author{Jens Raaby, Mads Hartmann \& Ulrik Rasmussen}

\begin{document}

\maketitle

\section*{Question 1}

% TODO: We probably need to make this more mathy
%       but the intuition should be sound

Given a graph $G$ a 1-tree bound is defined as follows

\begin{itemize}
  \item You select a vertex $v_1$ from $G$
  \item Remove all edges incident on $v_1$ from $G$ producing $G'$
  \item Find the minimum spanning tree $T$ of $G'$
  \item Add the two smallest edges $e_1$ and $e_2$ incident on $v_1$
    to $T$ producing a 1-tree $T_1$
\end{itemize}

We're asked to show that the 1-tree is a lower bound for TSP; that is,
that the weight of the Hamiltonian tour of G with the lowest weight is
equal to or greater than weight of $T_1$.

If we assume we have a Hamiltonian tour of $G$ that has has a lower
weight than the $T_1$ then it would require that there exists a
spanning tree of $G'$ that has a lower wight than $T$; this is
impossible since $T$ is a minimum spanning tree. Here's why: A
Hamiltonian tour of $G$ will also need to be a Hamiltonian path of
$G'$. In the best case the Hamiltonian path $H$ and the minimum
spanning tree $T$ will be the same. Since we're adding the two edges
$e_1, e_2$ with the lowest weight then the only way for $TSP$ to have
a lower weight than $T_1$ would be if the weight of $H$ was smaller
than $T$; a contradiction.


\section*{Question 2}
In this part we consider representing each monument as a circle, so
that when all circles are maximum size and do not overlap we have a
way of travelling to each monument by entering and leaving every
circle.

To form a linear program to maximise the radii of these circles, it is
a simple matter of representing each monument with a radius variable
and adding constraints such that any two circles to not overlap. In
the following $r_i$ denotes the radius of the circle placed on vertex
$i$, and $d(u,v)$ is the distance between vertices $u$ and $v$:

\begin{align*}
  \text{maximize} && \sum_{v\in V} r_{v} \\
  \text{subject to} && r_{u} + r_{v} \leq{}& d(u,v) & \forall (u,v) \in E, \\
                    && r_{v} \geq{}& 0 & \forall v \in V.
\end{align*}

\section*{Question 3}
How do you compute a lower bound from these circles?

Considering any single monument, its distance from any other monument
is at least the radius of that monument's circle. Since visiting all
the monuments requires both entering and leaving each of the circles,
a lower bound for a TSP problem could be expressed as $2 \sum_{v \in
  V} r_{v}$, in other words, double the sum of all the radii.  This
will be a fairly good lower bound if all the monuments are similar
distances apart. However, if many of the monuments are located in
small areas, then the radii will also be quite small. There may
therefore be a much larger distance between these groups of monuments,
meaning the actual TSP solution has a much higher cost. We can
mitigate this problem using the concept of \emph{moats} as described
in \cite{zones}, where clusters of nodes are chosen from which a
common moat can be grown until it hits a circle or another moat. We
then define the \emph{width} $w$ of a moat as the distance between its
perimeter and the perimeter of the circles that it encloses. Since a
valid TSP solution must enter and leave any cluster of nodes, we can
add $2w$ to the lower bound.

Assuming we have chosen some clustering of vertices, expressed as a
family of disjoint sets $C_i \in \mathcal{C}$ where $\bigcap_i C_i =
\emptyset$ and $\bigcup_i C_i = V$, a linear program maximizing
circles with moats can be formulated as follows.
\begin{align*}
  \text{maximize} && \sum_{v\in V} r_{v} + \sum_{C_i \in \mathcal{C}} w_i, \\
  \text{subject to} && r_{u} + r_{v} \leq{}& d(u,v) & \forall (u,v) \in E, u \in C_i, v \in C_i, \\
                    && r_u + r_v + w_i + w_j \leq{}& d(u,v) & \forall (u,v) \in E, u \in C_i, v \in C_j, i \not= j, \\
                    && r_{v} \geq{}& 0 & \forall v \in V.
\end{align*}

We note that the edges that are forced in a given branch in the
branch\&bound algorithm gives rise to a set of clusters.


\section*{Question 4}

We can use the mixed integer linear formulation by Miller, Tucker and
Zemlin to obtain a linear program of polynomial size with regards to
the number of nodes in the TSP problem. In the following, $d_{ij}$
denotes the distance from node $i$ to node $j$, and $x_{ij}$ is an
integer variable in the domain $\{0, 1\}$, where $x_{ij} = 1$ iff the
edge from node $i$ to $j$ is included in the solution. In addition,
there are $|V|-1$ auxiliary variables $u_i$ in the domain of the real
numbers, which are used to eliminate subtours:

\begin{align*}
  &\text{minimize} &&& \sum_{i\in V} \sum_{j \in V} d_{ij} x_{ij} \\
  &\text{subject to} &&& \sum_{j \in V} x_{ij} + \sum_{j \in V} x_{ji} ={}& 2 && \forall i \in V \\
  &                  &&& u_i - u_j + |V| \cdot x_{ij} \leq{}& |V| - 1 && \forall i,j \in V - \{1\}  \\
  &                  &&& u_i \geq{}& 0 && \forall i \in V - \{1\} \\
  &                  &&& x_{ij} \in{}& \{0, 1\} && \forall i,j \in V
\end{align*}

The first constraint ensures that each node has exactly one incoming
and one outgoing edge that is included in the solution. The second
constraint ensures that there is only a single tour in the solution,
namely the tour beginning and ending at node $1$.


\section*{Question 5}

We can relax the above mixed integer linear program by allowing the
edge variables $x_{ij}$ to take on real values. This yields the
following linear program, which does not include integer variables,
and hence can be solved efficiently:
\begin{align*}
  &\text{minimize} &&& \sum_{i\in V} \sum_{j \in V} d_{ij} x_{ij} \\
  &\text{subject to} &&& \sum_{j \in V} x_{ij} + \sum_{j \in V} x_{ji} ={}& 2 && \forall i \in V \\
  &                  &&& u_i - u_j + |V| \cdot x_{ij} \leq{}& |V| - 1 && \forall i,j \in V - \{1\}  \\
  &                  &&& u_i \geq{}& 0 && \forall i \in V - \{1\} \\
  &                  &&& x_{ij} \leq{}& 1 && \forall i,j \in V \\
  &                  &&& x_{ij} \geq{}& 0 && \forall i,j \in V
\end{align*}

The solution space for the original integer linear program is a subset
of the solution space for the relaxed version. Hence an optimal
solution to the relaxed program will be a lower bound for the integer
solution.


\section*{Question 6}

We have implemented the three lower bounds based on the Java code that
was handed out. The code archive that was turned in with this report
should be self-contained. To run the TSP solver with each of the lower
bounds, use the provided \texttt{Makefile}, by issuing \texttt{make
  \textit{bound}} on the terminal, where \emph{bound} is one of
``onetree'', ``radius'' or ``relax''.


For the three graphs, we get the following optimal values for TSP
tours:
\begin{align*}
  \mathit{OPT}_1 ={}& 8.649, \\
  \mathit{OPT}_2 ={}& 19.030, \\
  \mathit{OPT}_3 ={}& 26.753.
\end{align*}

We evaluate the following number of nodes for each lower bound. Note
that the number of nodes are not correlated with the actual time it
took to compute the solution, as some of these lower bounds has a
considerable overhead in our implementation, due to linear programs
being generated in inefficient ways. Thus, \emph{1-tree} completed
fastest, followed by \emph{relax}, followed by \emph{radius}.

\begin{center}
\begin{tabular}{r | c c c }

 & Instance 1 & Instance 2 & Instance 3 \\
\hline

Brute-force & 233902 & 1659236 & $\infty$ \\

1-Tree      & 2117 & 185 & 482435 \\

Radius      & 2283  & 586630 & $\infty$ \\

Relax       & 1377 & 90 & 70362 \\

\end{tabular}
\end{center}


\section*{Question 7}

\subsection{1-tree}

This lower bound method did not perform very well for the large graph
instance. The reason for this could be that the 1-tree does not give
any guarantees of the size of the cycle that it contains, just that it
\emph{will} contain a cycle. The 1-tree therefore will not necessarily
have any structure in common with the optimal solution, causing it to
underestimate the lower bound considerably.

We are not sure what can be done to improve this, but maybe a more
intelligent choice of the $1$-node will improve the lower bound.


\subsection{radius}

This lower bound method did not perform very well for neither the
medium sized or the large graph instance. It seems that the
performance of this lower bound depends a lot on the choice of
clusters used for calculating the moats.

We got a slight performance increase when ordering the edges in
ascending order according to their sizes, as this seems to form
clusters of nodes that are close to each other. This heuristic could
probably be improved by using a more advanced technique than sorting
the edges.


\subsection{relax}

This lower bound method performed fairly well with regards to the
number of nodes evaluated.

We are not sure how to improve this lower bound either, but one could
try to find more constraints to add to the linear program which would
still result in solutions being a lower bound for TSP, but which would
constrict the potential solution space.


\bibliographystyle{abbrv}
\bibliography{bibliography}
\end{document}
